### 2011-09-30 cover-letter.tex
final edits

\documentclass{letter}
\usepackage{graphics}
\usepackage{fullpage}
\topmargin=-0.2in
%\headheight=0in
\setlength\oddsidemargin{-0.2in}
\textheight=8.5in
\textwidth=7in
\def\apl#1#2{
\begin{letter}{#1}
\signature{Professor Dianne Cook\\Department of Statistics\\Iowa State University 
\\Ames, Iowa 50011-1210\\ Phone (515) 294 8865\\{\tt dicook@iastate.edu}}
\opening{{Dear #2}}

Please find attached our submission for the special issue of Environmetrics on Climate. The title is Glyph-maps for Visualizing Climate Data and Models, with authors Hadley Wickham, Heike Hofmann, Charlotte Wickham and myself. We hope that you find this paper interesting reading, and that it will make a valuable contribution  to the special issue.

\closing{Sincerely yours,
{\pdfimage width 1.5in {signature2.pdf}}}
\end{letter}}

\begin{document}

{{\pdfimage width 4in {letterhead.jpg}}}
\apl{Editors \\Environmetrics Special Issue on Climate}{Profs Guttorp, Sain and Wikle}

\end{document}